---
title: "Theoretical linear model SAMexplore"
author: "Milou Sep"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    toc: yes
    number_sections: yes
    df_print: kable
    fig_width: 10
editor_options: 
  chunk_output_type: console
---

# Preparations

The TM analysis of the SAM explore project.

First the required packages are added.
```{r setup, include=FALSE}
rm(list=ls())
# for imputed data
library(mice)
library(mitools)
#export
library(flextable) # To make publication ready tables for word, info: https://davidgohel.github.io/flextable/articles/overview.html
library(officer)
```

```{r data, include=FALSE}
# load imputed data, prepared for RFs / Boruta
source("R/SAMexplore_prepare.for.analyses.r")
```

transform lists to 'imputationLists'
```{r create implists}
imputationList(imp.nos)->imp.nos2
imputationList(imp.ims)->imp.ims2
imputationList(imp.des)->imp.des2
```


# Regression Models

## Variable selection
Variable selection for this model mostly based on the model described by [Kok etal. 2016](https://www.sciencedirect.com/science/article/pii/S0165032716304530?via%3Dihub),
where a step-wise regression is performed: 

* step1: relevant background variables  (not their interaction?)
* step2: a combined stress-exposure score (CTQ and LSC-r)
* step3: trait anxiety is added

Following this, for SAM explore:
* age, BMI
* Stress-exposure
* Trait anxiety
* AUCi sAA (d1)
* AUCi cort (d1)

The analysis is performed on imputed data (10 imputed datasets).

## Assumption check 
The assumptions for linear regression were checked on each imputed dataset, using "SAMexplore_assumptions_regression.RMD"
```{r eval=FALSE, include=FALSE}
# # Note: assumptions checked on
# with(imp.data, lm(eval(dependent_var)~1+age+BMI+stress_exposure*STAI_T_total*AUCi_sAA_1*AUCi_CORT_1))->M.full
# print(summary( pool(M.full)))
```


## Models & comparisons
Models were compared using the Wald test (and pool.compare / D1) https://rdrr.io/cran/mice/man/pool.compare.html

```{r formula with theoretical models}
theoretical_models <- function(dependent_var, imp.data, file.export, TableName){
  
  # # Full model
  with(imp.data, lm(eval(dependent_var)~1+
                      age+BMI+stress_exposure+STAI_T_total+AUCi_sAA_1+AUCi_CORT_1+ #main effects
                      stress_exposure:STAI_T_total+ # 2 way
                      stress_exposure:AUCi_sAA_1+
                      stress_exposure:AUCi_CORT_1+
                      STAI_T_total:AUCi_sAA_1+
                      STAI_T_total:AUCi_CORT_1 +
                      STAI_T_total:AUCi_sAA_1:AUCi_CORT_1+ 
                      stress_exposure:AUCi_sAA_1:AUCi_CORT_1
  ))->M.full
  
  # 3 way interaction (compared to full model): TA x sAA x Cort
  with(imp.data, lm(eval(dependent_var)~1+
                      age+BMI+stress_exposure+STAI_T_total+AUCi_sAA_1+AUCi_CORT_1+ #main effects
                      stress_exposure:STAI_T_total+ # 2 way
                      stress_exposure:AUCi_sAA_1+
                      stress_exposure:AUCi_CORT_1+
                      STAI_T_total:AUCi_sAA_1+
                      STAI_T_total:AUCi_CORT_1+
                      # STAI_T_total:AUCi_sAA_1:AUCi_CORT_1+ # 3way
                      stress_exposure:AUCi_sAA_1:AUCi_CORT_1
  ) )->M.no.TA.sAA.CORT 
  # Compare model without 3way to full model: effect of the 3way interaction
  D1(M.full,M.no.TA.sAA.CORT)->TA.sAA.CORT
  threeway.T.A.C<-c("Trait Anxiety x sAA x Cortisol", TA.sAA.CORT$result)
  
  # 3 way interaction (compared to full model): stress x sAA x cort
  with(imp.data, lm(eval(dependent_var)~1+
                      age+BMI+stress_exposure+STAI_T_total+AUCi_sAA_1+AUCi_CORT_1+ #main effects
                      stress_exposure:STAI_T_total+ # 2 way
                      stress_exposure:AUCi_sAA_1+
                      stress_exposure:AUCi_CORT_1+
                      STAI_T_total:AUCi_sAA_1+
                      STAI_T_total:AUCi_CORT_1+
                      STAI_T_total:AUCi_sAA_1:AUCi_CORT_1 #+ # 3way
                    # stress_exposure:AUCi_sAA_1:AUCi_CORT_1
  ))->M.no.St.sAA.CORT
  # Compare model without 3way to full model: effect of the 3way interaction
  D1(M.full,M.no.St.sAA.CORT)->St.sAA.CORT 
  threeway.S.A.C<-c("Life Adversity x sAA x Cortisol",St.sAA.CORT$result)
  
  # Model without 3-way interaction
  with(imp.data, lm(eval(dependent_var)~1+
                      age+BMI+stress_exposure+STAI_T_total+AUCi_sAA_1+AUCi_CORT_1+ #main effects
                      stress_exposure:STAI_T_total+ # 2 way
                      stress_exposure:AUCi_sAA_1+
                      stress_exposure:AUCi_CORT_1+
                      STAI_T_total:AUCi_sAA_1+
                      STAI_T_total:AUCi_CORT_1 #+
                    # STAI_T_total:AUCi_sAA_1:AUCi_CORT_1+ # 3way
                    # stress_exposure:AUCi_sAA_1:AUCi_CORT_1 +
  ))->M.no3way
  
  # 2 way effects (compared to model without 3way): TA x cort
  with(imp.data, lm(eval(dependent_var)~1+
                      age+BMI+stress_exposure+STAI_T_total+AUCi_sAA_1+AUCi_CORT_1+ #main effects
                      stress_exposure:STAI_T_total+ # 2 way
                      stress_exposure:AUCi_sAA_1+
                      stress_exposure:AUCi_CORT_1+
                      STAI_T_total:AUCi_sAA_1 #+
                    #  STAI_T_total:AUCi_CORT_1 #+
  ))->M.no.TA.CORT
  # Compare model without 2way to model without 3way interaction: effect of the 2way interaction
  D1(M.no3way,M.no.TA.CORT)->TA.CORT 
  twoway.T.C<-c("Trait Anxiety x Cortisol",TA.CORT$result)
  
  # 2 way effects (compared to model without 3way): TA x sAA
  with(imp.data, lm(eval(dependent_var)~1+
                      age+BMI+stress_exposure+STAI_T_total+AUCi_sAA_1+AUCi_CORT_1+ #main effects
                      stress_exposure:STAI_T_total+ # 2 way
                      stress_exposure:AUCi_sAA_1+
                      stress_exposure:AUCi_CORT_1+
                      #  STAI_T_total:AUCi_sAA_1 #+
                      STAI_T_total:AUCi_CORT_1
  ))->M.no.TA.sAA
  # Compare model without 2way to model without 3way interaction: effect of the 2way interaction
  D1(M.no3way,M.no.TA.sAA)->TA.sAA
  twoway.T.A<-c("Trait Anxiety x sAA",TA.sAA$result)
  
  # 2 way effects (compared to model without 3way): Stress x cort
  with(imp.data, lm(eval(dependent_var)~1+
                      age+BMI+stress_exposure+STAI_T_total+AUCi_sAA_1+AUCi_CORT_1+ #main effects
                      stress_exposure:STAI_T_total+ # 2 way
                      stress_exposure:AUCi_sAA_1+
                      #  stress_exposure:AUCi_CORT_1+
                      STAI_T_total:AUCi_sAA_1 +
                      STAI_T_total:AUCi_CORT_1
  ))->M.no.St.CORT
  # Compare model without 2way to model without 3way interaction: effect of the 2way interaction
  D1(M.no3way,M.no.St.CORT)->St.CORT
  twoway.S.C<-c("Life Adversity x Cortisol", St.CORT$result)
  
  # 2 way effects (compared to model without 3way): stress x sAA
  with(imp.data, lm(eval(dependent_var)~1+
                      age+BMI+stress_exposure+STAI_T_total+AUCi_sAA_1+AUCi_CORT_1+ #main effects
                      stress_exposure:STAI_T_total+ # 2 way
                      # stress_exposure:AUCi_sAA_1+
                      stress_exposure:AUCi_CORT_1+
                      STAI_T_total:AUCi_sAA_1 +
                      STAI_T_total:AUCi_CORT_1
  ))->M.no.St.sAA
  # Compare model without 2way to model without 3way interaction: effect of the 2way interaction
  D1(M.no3way,M.no.St.sAA)->St.sAA
  twoway.S.A<-c("Life Adversity x sAA", St.sAA$result)
  
  # 2 way effects (compared to model without 3way): stress x TA 
  with(imp.data, lm(eval(dependent_var)~1+
                      age+BMI+stress_exposure+STAI_T_total+AUCi_sAA_1+AUCi_CORT_1+ #main effects
                      # stress_exposure:STAI_T_total+ # 2 way
                      stress_exposure:AUCi_sAA_1+
                      stress_exposure:AUCi_CORT_1+
                      STAI_T_total:AUCi_sAA_1 +
                      STAI_T_total:AUCi_CORT_1 #+
  ))->M.no.St.TA
  # Compare model without 2way to model without 3way interaction: effect of the 2way interaction
  D1(M.no3way,M.no.St.TA)->St.TA
  twoway.S.T<-c("Life Adversity x Trait Anxiety", St.TA$result)
  
  # Model without 2 way interactions (= only main effects) 
  with(imp.data, lm(eval(dependent_var)~1+age+BMI+stress_exposure+STAI_T_total+AUCi_sAA_1+AUCi_CORT_1#+ #main effects
  ))->M.no2way 
  
  # Main effect: CORT
  with(imp.data, lm(eval(dependent_var)~1+age+BMI+
                      stress_exposure+
                      STAI_T_total+
                      AUCi_sAA_1#+
                    # AUCi_CORT_1#+ #main effects
  ))->M.Cort
  # Compare model without variable to model without 2way interaction: main effect of variable
  D1(M.no2way,M.Cort)->mCORT
  main.C<-c("Cortisol", mCORT$result)
  
  # Main effect: sAA
  with(imp.data, lm(eval(dependent_var)~1+age+BMI+
                      stress_exposure+
                      STAI_T_total+
                      # AUCi_sAA_1#+
                      AUCi_CORT_1#+ #main effects
  ))->M.sAA
  # Compare model without variable to model without 2way interaction: main effect of variable
  D1(M.no2way,M.sAA)->msAA
  main.A<-c("sAA", msAA$result)
  
  # Main effect: TA
  with(imp.data, lm(eval(dependent_var)~1+age+BMI+
                      stress_exposure+
                      # STAI_T_total+
                      AUCi_sAA_1+
                      AUCi_CORT_1#+ #main effects
  ))->M.TA
  # Compare model without variable to model without 2way interaction: main effect of variable
  D1(M.no2way,M.TA)->mTA
  main.T<-c("Trait Anxiety", mTA$result)
  
  # Main effect: stress
  with(imp.data, lm(eval(dependent_var)~1+age+BMI+
                      #stress_exposure+
                      STAI_T_total+
                      AUCi_sAA_1+
                      AUCi_CORT_1#+ #main effects
  ))->M.ST
  # Compare model without variable to model without 2way interaction: main effect of variable
  D1(M.no2way,M.ST)->mST
  main.S<-c("Life Adversity", mST$result)
  
  # # Collect results of model comparisons in vectors
  results <- data.frame(rbind(
    # hypothesized interaction in TM
    threeway.T.A.C,
    threeway.S.A.C,
    twoway.S.A,
    twoway.S.C,
    twoway.T.A,
    twoway.T.C,
    main.A,
    main.C ,
    # additional interactions
    twoway.S.T,
    main.S,
    main.T))
  
  colnames(results) <- c( "Model",  "Wald statistic (D1)", "df1",  "df2",    "P(>F)", "RIV")
  results %>% mutate_at(. ,c(2:6), as.character) %>% mutate_at(. ,c(2:6), as.numeric) ->results
  
  # Use flextable layout to prepare table (for word)
  flextable(results) %>% 
    colformat_num(x=.,
                  j = c(2,4,5,6),
                  big.mark = "",
                  digits = 2) %>% 
    colformat_num(x=.,j = c( "df1"),big.mark = "",digits = 0) %>%
    set_caption(., TableName, html_escape = T)%>%
    theme_vanilla(.)%>%
    bold(., i= ~`P(>F)` < 0.07 , j="P(>F)") %>%
    autofit(.)->results.table
  
  # if you want individual tables to word
  if (file.export == T){
    doc <- read_docx()
    doc <- body_add_flextable(doc, value = results.table, align="center")
    print(doc, target = paste0("results/",TableName, date(),".docx"))
  }
  
  return(results.table)  
}

```


# No Stress group
## Results
- 3-way interaction between trait anxiety * sAA (learning) * cortisol (learning) on emotional memory

```{r call function no stress, echo=FALSE}
theoretical_models(sym("emotional.memory"), imp.nos2, F,"No Stress - Emotional MC" ) ->emo_no
emo_no
theoretical_models(sym("neutral.memory"), imp.nos2, F,"No Stress - Neutral MC" ) ->neu_no
neu_no
theoretical_models(sym("fear.memory"), imp.nos2 , F,"No Stress - Fearful MC") ->feM_no
feM_no
theoretical_models(sym("fear.learning"), imp.nos2 , F,"No Stress - Fearful Learning") ->feL_no
feL_no
```


# Acute phase stress response group
## Results
- Main effect cumulative stress-exposure during life on emotional memory contextualization immediately after stress
- Main effect cortisol day1 (learning) on neutral memory contextualization immediately after stress
- Two way interaction stress-exposure x sAA (learning) and trait anxiety x cortisol (learning) on Fearful MC after stress
- no effects on Fearful learning after stress

```{r call function Acute phase stress response, echo=FALSE}
theoretical_models(sym("emotional.memory"), imp.ims2 , F,"Acute phase stress response - Emotional MC") ->emo_im
emo_im
theoretical_models(sym("neutral.memory"), imp.ims2, F,"Acute phase stress response - Neutral MC" ) ->neu_im
neu_im
theoretical_models(sym("fear.memory"), imp.ims2 , F,"Acute phase stress response - Fearful MC") ->feM_im
feM_im
theoretical_models(sym("fear.learning"), imp.ims2, F,"Acute phase stress response - Fearful Learning" ) ->feL_im
feL_im
```


# Recovery phase stress responses
## Results
- 2h after stress, no effect on emotional memory, Fearful memory or Fearful learning
- Two way interaction stress exposure x sAA (learning) and stress exposure x Cortisol (learning) on neutral memory contextualization

```{r call function Recovery phase stress responses, echo=FALSE}
theoretical_models(sym("emotional.memory"), imp.des2, F,"Recovery phase stress responses - Emotional MC" ) ->emo_de
emo_de
theoretical_models(sym("neutral.memory"), imp.des2, F,"Recovery phase stress responses - Neutral MC" ) ->neu_de
neu_de
theoretical_models(sym("fear.memory"), imp.des2 , F,"Recovery phase stress responses - Fearful MC") ->feM_de
feM_de
theoretical_models(sym("fear.learning"), imp.des2, F,"Recovery phase stress responses - Fearful Learning" ) ->feL_de
feL_de
```

>  RIV: relative increase in variance due to nonresponse  https://rdrr.io/cran/mitml/man/testModels.html
can be interpreted as the proportion of the variation attributable to the missing data [https://stefvanbuuren.name/fimd/sec-whyandwhen.html]

# Combine for Word-export

```{r word export}
word_export <- read_docx()
body_add_flextable(word_export, neu_no)
body_add_par(word_export, value = "")
body_add_flextable(word_export, emo_no)
body_add_par(word_export, value = "")
body_add_flextable(word_export, feM_no)
# body_add_par(word_export, value = "")
# body_add_flextable(word_export, feL_no)

body_add_par(word_export, value = "")
body_add_flextable(word_export, neu_im)
body_add_par(word_export, value = "")
body_add_flextable(word_export, emo_im)
body_add_par(word_export, value = "")
body_add_flextable(word_export, feM_im)

body_add_par(word_export, value = "")
body_add_flextable(word_export, neu_de)
body_add_par(word_export, value = "")
body_add_flextable(word_export, emo_de)
body_add_par(word_export, value = "")
body_add_flextable(word_export, feM_de)

print(word_export, target = paste0("results/TM_statistics.2", date(),".docx"))
```
