---
title: "Model performance evaluation: TM, RF, TM+RF"
author: "Milou Sep"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---


## Required packages and (preprocessed and imputed) data
```{r setup}
rm(list=ls())
library(caret)
library(DALEX) # for model performance measures. https://pbiecek.github.io/ema/modelPerformance.html#modelPerformanceR
library(ggplot2)
library(ggpubr) 
library(patchwork)
```
load imputed data, prepared for SAM_explore analysis RFs / Boruta / lm's
```{r data, results='hide', message=FALSE}
source("R/SAMexplore_prepare.for.analyses.r")
seed=3820
# rf settings
m.try=2
n.tree=1500
```


# Model performance evaluation function

Model performance evaluated via R2 and RMSE (https://stats.stackexchange.com/questions/56302/what-are-good-rmse-values)
- RMSE has the same unit as the dependent variable (DV)
- "there is no such thing as a good RMSE, because it is scale-dependent, i.e. dependent on your dependent variable. Hence one can not claim a universal number as a good RMSE."

```{r}
model.performance<-function(seed, method, model, dataset, preprocessing.formula, dependent.var){
  # process data
  dataset2<-preprocessing.formula(dataset)
  # Setup cross-validation
  control <- trainControl(method="repeatedcv", number=5, repeats=10)
  # train model according to method
  if (method == 'lm'){
    set.seed(seed)
    train(eval(model),method=method,data=dataset2, trControl=control)->trained.model
  }else if (method == 'rf'){
    set.seed(seed)
    train(eval(model), dataset2, method=method, trControl=control, ntree=n.tree)->trained.model
  }
  set.seed(seed)
  # calculate model performance
  DALEX::explain(trained.model,
                 data=dataset2, # performance calculated on this dataset
                 y=dataset2[,dependent.var]
  )->explained.model
  
  set.seed(seed)
  DALEX::model_performance(explained.model)->performance

  performance$measures$rmse->rmse.value
  performance$measures$r2->r2.value
  
  #collect results
  out<-list(explained.model, r2.value, rmse.value)
  return(out)
}
```

# theoretical linear models
- information on the use of linear model & predict() together, see https://www.pluralsight.com/guides/linear-lasso-and-ridge-regression-with-r/ 
- models with covariates (age & BMI) and the significant term(s) per memory type / group are evaluated

### no stress group
- emotional memory: 3-way interaction TA x cort x sAA
```{r}
lapply(1:M,function(x){
  model.performance(seed, method='lm', model=expression(emotional.memory~1+age+BMI+
                                                          STAI_T_total:AUCi_sAA_1:AUCi_CORT_1), 
                    dataset=imp.nos[[x]], prep.emo,"emotional.memory")})->t.ne
# collect r2's
lapply(1:M,function(x){t.ne[[x]][[2]]}) %>% unlist ->t.r2.ne
# collect rmse's
lapply(1:M,function(x){t.ne[[x]][[3]]}) %>% unlist ->t.rmse.ne
```

### acute phase group
- fear memory: interactions:
  - trait anxiety x cortisol (i,day1) and 
  - stress-exposure x sAA (i,day1)
info on code for interaction terms: https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/formula
```{r}
lapply(1:M,function(x){
  model.performance(seed, method='lm', model=expression(fear.memory~1+age+BMI+
                                                          STAI_T_total:AUCi_CORT_1+
                                                          stress_exposure:AUCi_sAA_1), 
                    dataset=imp.ims[[x]], prep.fear.m,"fear.memory")})->t.if
# collect r2's
lapply(1:M,function(x){t.if[[x]][[2]]}) %>% unlist ->t.r2.if
# collect rmse's
lapply(1:M,function(x){t.if[[x]][[3]]}) %>% unlist ->t.rmse.if
```

- emotional: main effect cumulative stress on emotional memory
```{r}
lapply(1:M,function(x){
  model.performance(seed, method='lm', model=expression(emotional.memory~1+age+BMI+
                                                          stress_exposure), 
                    dataset=imp.ims[[x]], prep.emo,"emotional.memory")})->t.ie
# collect r2's
lapply(1:M,function(x){t.ie[[x]][[2]]}) %>% unlist ->t.r2.ie
# collect rmse's
lapply(1:M,function(x){t.ie[[x]][[3]]}) %>% unlist ->t.rmse.ie
```

- neutral: main effect AUCi cortisol d1 on neutral memory
```{r}
lapply(1:M,function(x){
  model.performance(seed, method='lm', model=expression(neutral.memory~1+age+BMI+
                                                          AUCi_CORT_1), 
                    dataset=imp.ims[[x]], prep.neu,"neutral.memory")})->t.in
# collect r2's
lapply(1:M,function(x){t.in[[x]][[2]]}) %>% unlist ->t.r2.in
# collect rmse's
lapply(1:M,function(x){t.in[[x]][[3]]}) %>% unlist ->t.rmse.in
```

### recovery phase group
- no effects on emotional or fear memory
- neutral memory: interactions:
  - stress-exposure x cortisol (i,day1) and
  - stress-exposure x sAA (i,day1)
```{r}
lapply(1:M,function(x){
  model.performance(seed, method='lm', model=expression(neutral.memory~1+age+BMI+
                                                          stress_exposure:AUCi_sAA_1+
                                                          stress_exposure:AUCi_CORT_1), 
                    dataset=imp.des[[x]], prep.neu,"neutral.memory")})->t.dn
# collect r2's
lapply(1:M,function(x){t.dn[[x]][[2]]}) %>% unlist ->t.r2.dn
# collect rmse's
lapply(1:M,function(x){t.dn[[x]][[3]]}) %>% unlist ->t.rmse.dn
```


# Boruta RF models

- Follow up boruta findings with Random-forests per imputed dataset
- good manual for RF after Boruta (https://www.kaggle.com/cdupin/using-boruta-feature-selection-wtih-random-forest)

### no stress
- fear memory: HEXACO_hh.mean, AUCi_stais_2
```{r}
lapply(1:M,function(x){
  model.performance(seed, method='rf',
                    model=expression(fear.memory~HEXACO_hh.mean+AUCi_stais_2),
                    dataset=imp.nos[[x]],
                    prep.fear.m,"fear.memory")})->rf.nf
# collect r2's
lapply(1:M,function(x){rf.nf[[x]][[2]]}) %>% unlist ->rf.r2.nf
# collect rmse's
lapply(1:M,function(x){rf.nf[[x]][[3]]}) %>% unlist ->rf.rmse.nf
```

- emotional: HEXACO_ag.mean, AUCi_panas_PosA_1
```{r}
lapply(1:M,function(x){
  model.performance(seed, method='rf',
                    model=expression(emotional.memory~HEXACO_ag.mean+AUCi_panas_PosA_1),
                    dataset=imp.nos[[x]],
                    prep.emo, "emotional.memory")})->rf.ne
# collect r2's
lapply(1:M,function(x){rf.ne[[x]][[2]]}) %>% unlist ->rf.r2.ne
# collect rmse's
lapply(1:M,function(x){rf.ne[[x]][[3]]}) %>% unlist ->rf.rmse.ne
```

- neutral: HEXACO_co.mean, VTCI_CO_total
```{r}
lapply(1:M,function(x){
  model.performance(seed, method='rf', 
                    model=expression(neutral.memory~HEXACO_co.mean+VTCI_CO_total),
                    dataset=imp.nos[[x]],
                    prep.neu, "neutral.memory")})->rf.nn
# collect r2's
lapply(1:M,function(x){rf.nn[[x]][[2]]}) %>% unlist ->rf.r2.nn
# collect rmse's
lapply(1:M,function(x){rf.nn[[x]][[3]]}) %>% unlist ->rf.rmse.nn
```


### acute phase group
- fear memory: JTV_total HEXACO_extr.mean identified by Boruta, but NS in permutation tests (model performance not evaluated)

- emotional memory: JTV_total HEXACO_hh.mean AUCi_VAS_mood_2
```{r}
lapply(1:M,function(x){
  model.performance(seed, method='rf',
                    model=expression(emotional.memory~JTV_total+HEXACO_hh.mean+AUCi_VAS_mood_2),
                    dataset=imp.ims[[x]],
                    prep.emo, "emotional.memory")})->rf.ie
# collect r2's
lapply(1:M,function(x){rf.ie[[x]][[2]]}) %>% unlist ->rf.r2.ie
# collect rmse's
lapply(1:M,function(x){rf.ie[[x]][[3]]}) %>% unlist ->rf.rmse.ie
```

- neutral memory: AUCi_sAA_2, stress_exposure
```{r}
lapply(1:M,function(x){
  model.performance(seed, method='rf',
                    model=expression(neutral.memory~AUCi_sAA_2+stress_exposure),
                    dataset=imp.ims[[x]],
                    prep.neu, "neutral.memory")})->rf.in
# collect r2's
lapply(1:M,function(x){rf.in[[x]][[2]]}) %>% unlist ->rf.r2.in
# collect rmse's
lapply(1:M,function(x){rf.in[[x]][[3]]}) %>% unlist ->rf.rmse.in
```

### recovery phase group
- fear memory: non
- emotional memory:AUCg_sAA_2 AUCg_VAS_anxiety_2 AUCi_VAS_mood_2
```{r}
lapply(1:M,function(x){
  model.performance(seed, method='rf',
                    model=expression(emotional.memory~AUCg_sAA_2+AUCg_VAS_anxiety_2+AUCi_VAS_mood_2),
                    dataset=imp.des[[x]],
                    prep.emo, "emotional.memory")})->rf.de
# collect r2's
lapply(1:M,function(x){rf.de[[x]][[2]]}) %>% unlist ->rf.r2.de
# collect rmse's
lapply(1:M,function(x){rf.de[[x]][[3]]}) %>% unlist ->rf.rmse.de
```

- neutral memory: age AUCi_sAA_1 AUCi_CORT_2 AUCg_VAS_mood_1 AUCi_stais_1
```{r}
lapply(1:M,function(x){
  model.performance(seed, method='rf',
                    model=expression(neutral.memory~age+AUCi_sAA_1+AUCi_CORT_2+AUCg_VAS_mood_1+AUCi_stais_1),
                    dataset=imp.des[[x]],
                    prep.neu, "neutral.memory")})->rf.dn
# collect r2's
lapply(1:M,function(x){rf.dn[[x]][[2]]}) %>% unlist ->rf.r2.dn
# collect rmse's
lapply(1:M,function(x){rf.dn[[x]][[3]]}) %>% unlist ->rf.rmse.dn
```


# Ensemble models: TM + RF
- linear and non-linear terms in linear models (based on plots from boruta variables) (https://stats.stackexchange.com/questions/25975/how-to-add-second-order-terms-into-the-model-in-r)
- Note. the polynomial terms are based on plot inspections

### no stress group: predictors of context-dependency without acute stress

####  neutral memory without stress

- TM:
+ non
- RF: 
+ HEXACO_co.mean ^ 3
+ VTCI_CO_total

```{r}
lapply(1:M,function(x){
  model.performance(seed, method='lm', model=expression(neutral.memory~1+
                                                          # RF
                                                          poly(HEXACO_co.mean,3)+
                                                          VTCI_CO_total),
                    dataset=imp.nos[[x]], prep.neu,"neutral.memory")})->ens.nn
# collect r2's
lapply(1:M,function(x){ens.nn[[x]][[2]]}) %>% unlist -> ens.r2.nn
# collect rmse's
lapply(1:M,function(x){ens.nn[[x]][[3]]}) %>% unlist -> ens.rmse.nn
```

#### emotional memory without stress

- TM:
+ Trait Anxiety x sAA x Cortisol
- RF: 
+ HEXACO_ag.mean
+ AUCi_panas_PosA_1 ^ 2 

```{r}
lapply(1:M,function(x){
  model.performance(seed, method='lm', model=expression(emotional.memory~1+
                                                          # TM
                                                          STAI_T_total:AUCi_sAA_1:AUCi_CORT_1 +
                                                          # RF
                                                          HEXACO_ag.mean+
                                                          poly(AUCi_panas_PosA_1,2)),
                    dataset=imp.nos[[x]], prep.emo,"emotional.memory")})->ens.ne
# collect r2's
lapply(1:M,function(x){ens.ne[[x]][[2]]}) %>% unlist -> ens.r2.ne
# collect rmse's
lapply(1:M,function(x){ens.ne[[x]][[3]]}) %>% unlist -> ens.rmse.ne
```

#### fear memory without acute stress

- TM:
+ no predictors
- RF: 
+ HEXACO_hh.mean ^ 3
+ AUCi_stais_2

```{r}
lapply(1:M,function(x){
  model.performance(seed, method='lm', model=expression(fear.memory~1+
                                                          # RF
                                                          poly(HEXACO_hh.mean,3)+
                                                          AUCi_stais_2),
                    dataset=imp.nos[[x]], prep.fear.m,"fear.memory")})->ens.nf
# collect r2's
lapply(1:M,function(x){ens.nf[[x]][[2]]}) %>% unlist -> ens.r2.nf
# collect rmse's
lapply(1:M,function(x){ens.nf[[x]][[3]]}) %>% unlist -> ens.rmse.nf
```

### acute phase group: predictors of context-dependency ~ 30min after acute stress

#### neutral memory ~30 min after acute stress

- TM: 
+ cortisol (auci, d1)
- RF: 
+ AUCi_sAA_2
+ stress_exposure

```{r}
lapply(1:M,function(x){
  model.performance(seed, method='lm', model=expression(neutral.memory~1+
                                                          # TM
                                                          AUCi_CORT_1+
                                                          # RF
                                                          AUCi_sAA_2+ 
                                                          stress_exposure), 
                    dataset=imp.ims[[x]], prep.neu,"neutral.memory")})->ens.in
# collect r2's
lapply(1:M,function(x){ens.in[[x]][[2]]}) %>% unlist -> ens.r2.in
# collect rmse's
lapply(1:M,function(x){ens.in[[x]][[3]]}) %>% unlist -> ens.rmse.in
```

#### emotional memory ~30 min after acute stress

- TM:
+ [life-time stress, Note, stress_exposure from theoretical model is replaced by JTV_total (from Boruta)]
- RF:
+ JTV_total
+ HEXACO_hh.mean
+ AUCi_VAS_mood_2

```{r}
lapply(1:M,function(x){
  model.performance(seed, method='lm', model=expression(emotional.memory~1+
                                                          #RF (& TM)
                                                          JTV_total+ 
                                                          #RF
                                                          HEXACO_hh.mean+
                                                          AUCi_VAS_mood_2),
                    dataset=imp.ims[[x]], prep.emo,"emotional.memory")})->ens.ie
# collect r2's
lapply(1:M,function(x){ens.ie[[x]][[2]]}) %>% unlist -> ens.r2.ie
# collect rmse's
lapply(1:M,function(x){ens.ie[[x]][[3]]}) %>% unlist -> ens.rmse.ie
```

#### fear memory ~30 min after acute stress

- TM:
+ STAI_T_total:AUCi_CORT_1
+ stress_exposure:AUCi_sAA_1

- RF:  JTV_total & HEXACO_extr.mean Boruta selected, but model NS in permutation tests (so not included in model)

```{r}
lapply(1:M,function(x){
  model.performance(seed, method='lm', model=expression(fear.memory~1+
                                                          #TM
                                                          STAI_T_total:AUCi_CORT_1+
                                                          #TM
                                                          stress_exposure:AUCi_sAA_1
                                                          ), 
                    dataset=imp.ims[[x]], prep.fear.m,"fear.memory")})->ens.if
# collect r2's
lapply(1:M,function(x){ens.if[[x]][[2]]}) %>% unlist -> ens.r2.if
# collect rmse's
lapply(1:M,function(x){ens.if[[x]][[3]]}) %>% unlist -> ens.rmse.if
```


### recovery phase group: predictors of context-dependency ~2h after acute stress

#### neutral memory ~2h after acute stress

- TM:
+ Stress x sAA
+ Stress x Cortisol
- RF: 
+ age
+ AUCi_sAA_1   [note (AUCi_sAA_1 not separately included because present in interaction TM)]
+ AUCi_CORT_2
+ AUCg_VAS_mood_1
+ AUCi_stais_1^2

```{r}
lapply(1:M,function(x){
  model.performance(seed, method='lm', model=expression(neutral.memory~1+
                                                          #TM
                                                          stress_exposure:AUCi_CORT_1+
                                                          stress_exposure:AUCi_sAA_1+
                                                          #RF (AUCi_sAA_1 not separately included because present in interaction TM)
                                                          age+
                                                          AUCi_CORT_2+
                                                          AUCg_VAS_mood_1+ 
                                                          poly(AUCi_stais_1,2)),
                    dataset=imp.des[[x]], prep.neu,"neutral.memory")})->ens.dn
# collect r2's
lapply(1:M,function(x){ens.dn[[x]][[2]]}) %>% unlist -> ens.r2.dn
# collect rmse's
lapply(1:M,function(x){ens.dn[[x]][[3]]}) %>% unlist -> ens.rmse.dn
```

#### emotional memory ~2h after acute stress

- TM: 
+ no predictors
- RF: 
+ AUCg_sAA_2 ^ 2
+ AUCg_VAS_anxiety_2 ^2
+ AUCi_VAS_mood_2

```{r}
lapply(1:M,function(x){
  model.performance(seed, method='lm', model=expression(emotional.memory~1+
                                                          #RF
                                                          poly(AUCg_sAA_2,2)+
                                                          poly(AUCg_VAS_anxiety_2,2)+
                                                          AUCi_VAS_mood_2), 
                    dataset=imp.des[[x]], prep.emo,"emotional.memory")})->ens.de
# collect r2's
lapply(1:M,function(x){ens.de[[x]][[2]]}) %>% unlist -> ens.r2.de
# collect rmse's
lapply(1:M,function(x){ens.de[[x]][[3]]}) %>% unlist -> ens.rmse.de
```

#### fear memory ~2h after acute stress

- TM:
+ no predictors
- RF:
+ no variables


# Merge performance measures

### R2: calculate mean R2
```{r}
# results matrix
r2.table<-data.frame(matrix(nrow=23,ncol=5))
colnames(r2.table)<-c('analysis', 'group','memory','m.r2',"sd.r2")

# results predictions random forest
r2.table[1,]<-c('RF', "N",'F', mean(rf.r2.nf), sd(rf.r2.nf))
r2.table[2,]<-c('RF', "I",'F', NA, NA) # NS in permutation tests
r2.table[3,]<-c('RF', "N",'E', mean(rf.r2.ne), sd(rf.r2.ne))
r2.table[4,]<-c('RF', "I",'E', mean(rf.r2.ie), sd(rf.r2.ie))
r2.table[5,]<-c('RF', "D",'E', mean(rf.r2.de), sd(rf.r2.de))
r2.table[6,]<-c('RF', "N",'N', mean(rf.r2.nn), sd(rf.r2.nn))
r2.table[7,]<-c('RF', "I",'N', mean(rf.r2.in), sd(rf.r2.in))
r2.table[8,]<-c('RF', "D",'N', mean(rf.r2.dn), sd(rf.r2.dn))

# results predictions linear model
r2.table[9,]<-c('LM', "N",'E', mean(t.r2.ne), sd(t.r2.ne))
r2.table[10,]<-c('LM', "I",'F', mean(t.r2.if), sd(t.r2.if))
r2.table[11,]<-c('LM', "I",'E', mean(t.r2.ie), sd(t.r2.ie))
r2.table[12,]<-c('LM', "I",'N', mean(t.r2.in), sd(t.r2.in))
r2.table[13,]<-c('LM', "D",'N', mean(t.r2.dn), sd(t.r2.dn))

# results predictions ensemble model: Linear models with polynomial terms
r2.table[14,]<-c('ensemble', "N",'F', mean(ens.r2.nf), sd(ens.r2.nf))
r2.table[15,]<-c('ensemble', "I",'F', mean(ens.r2.if), sd(ens.r2.if))
r2.table[16,]<-c('ensemble', "N",'E', mean(ens.r2.ne), sd(ens.r2.ne))
r2.table[17,]<-c('ensemble', "I",'E', mean(ens.r2.ie), sd(ens.r2.ie))
r2.table[18,]<-c('ensemble', "D",'E', mean(ens.r2.de), sd(ens.r2.de))
r2.table[19,]<-c('ensemble', "N",'N', mean(ens.r2.nn), sd(ens.r2.nn))
r2.table[20,]<-c('ensemble', "I",'N', mean(ens.r2.in), sd(ens.r2.in))
r2.table[21,]<-c('ensemble', "D",'N', mean(ens.r2.dn), sd(ens.r2.dn))
# empty lines, better for layout plots
r2.table[22,]<-c('LM', "N",'N', NA,NA) 
r2.table[23,]<-c('LM', "D",'F', NA,NA)

saveRDS(r2.table,'results/r2.table.5cv.RDS')
readRDS('results/r2.table.5cv.RDS')->r2.table
```

### RMSE: calculate mean rsme
```{r}
# results matrix
rmse.table<-data.frame(matrix(nrow=23,ncol=5))
colnames(rmse.table)<-c('analysis', 'group','memory','m.rmse',"sd.rmse")

# results predictions random forest
rmse.table[1,]<-c('RF', "N",'F', mean(rf.rmse.nf), sd(rf.rmse.nf))
rmse.table[2,]<-c('RF', "I",'F', NA,NA) # NS in permutation tests
rmse.table[3,]<-c('RF', "N",'E', mean(rf.rmse.ne), sd(rf.rmse.ne))
rmse.table[4,]<-c('RF', "I",'E', mean(rf.rmse.ie), sd(rf.rmse.ie))
rmse.table[5,]<-c('RF', "D",'E', mean(rf.rmse.de), sd(rf.rmse.de))
rmse.table[6,]<-c('RF', "N",'N', mean(rf.rmse.nn), sd(rf.rmse.nn))
rmse.table[7,]<-c('RF', "I",'N', mean(rf.rmse.in), sd(rf.rmse.in))
rmse.table[8,]<-c('RF', "D",'N', mean(rf.rmse.dn), sd(rf.rmse.dn))

# results predictions linear model
rmse.table[9,]<-c('LM', "N",'E', mean(t.rmse.ne), sd(t.rmse.ne))
rmse.table[10,]<-c('LM', "I",'F', mean(t.rmse.if), sd(t.rmse.if))
rmse.table[11,]<-c('LM', "I",'E', mean(t.rmse.ie), sd(t.rmse.ie))
rmse.table[12,]<-c('LM', "I",'N', mean(t.rmse.in), sd(t.rmse.in))
rmse.table[13,]<-c('LM', "D",'N', mean(t.rmse.dn), sd(t.rmse.dn))

# results predictions ensemble model: Linear models with polynomial terms
rmse.table[14,]<-c('ensemble', "N",'F', mean(ens.rmse.nf), sd(ens.rmse.nf))
rmse.table[15,]<-c('ensemble', "I",'F', mean(ens.rmse.if), sd(ens.rmse.if))
rmse.table[16,]<-c('ensemble', "N",'E', mean(ens.rmse.ne), sd(ens.rmse.ne))
rmse.table[17,]<-c('ensemble', "I",'E', mean(ens.rmse.ie), sd(ens.rmse.ie))
rmse.table[18,]<-c('ensemble', "D",'E', mean(ens.rmse.de), sd(ens.rmse.de))
rmse.table[19,]<-c('ensemble', "N",'N', mean(ens.rmse.nn), sd(ens.rmse.nn))
rmse.table[20,]<-c('ensemble', "I",'N', mean(ens.rmse.in), sd(ens.rmse.in))
rmse.table[21,]<-c('ensemble', "D",'N', mean(ens.rmse.dn), sd(ens.rmse.dn))
# empty lines, better for layout plots
rmse.table[22,]<-c('LM', "N",'N', NA,NA) 
rmse.table[23,]<-c('LM', "D",'F', NA,NA)

saveRDS(rmse.table,'results/rmse.table.5cv.RDS')
readRDS('results/rmse.table.5cv.RDS')->rmse.table
```

# Visualize model performance 
of Linear models, random forest, and ensamble models
```{r}
# merge dataframes with goodness of fit measures (R2 and RMSE)
full_join(r2.table,rmse.table, by=c("analysis", "group", "memory"))->merged.model.performance
# change factor names and order
merged.model.performance$analysis = factor(merged.model.performance$analysis, levels=c("LM","RF","ensemble"))
merged.model.performance$memory = factor(merged.model.performance$memory, levels=c("F","E","N"), labels = c("Fearful","Emotional","Neutral"))
merged.model.performance$group = factor(merged.model.performance$group, levels=c("I","D","N"), labels = c("Acute phase","Recovery phase","No stress"))
# change variables to numeric
merged.model.performance %>% mutate(across(c("m.r2", "sd.r2", "m.rmse", "sd.rmse"), as.numeric))->merged.model.performance

# create a list with plot layout specifications
plot.list<- list(
  facet_grid(group ~ memory),
  theme_minimal(),
  rotate_x_text(angle = 45),
  labs(x="")
)

# create R2 plot 
merged.model.performance %>% 
  ggplot(aes(x=analysis, y=(m.r2)))+
  geom_bar(stat = "identity")+
  geom_errorbar(aes(ymin=(m.r2)-(sd.r2),
                    ymax=(m.r2)+(sd.r2))) +
  plot.list +
  labs(y= expression(paste('averaged ',R^{2})))->p.r2
p.r2

# create RMSE Plot
merged.model.performance %>% 
  ggplot(aes(x=analysis, y=(m.rmse)))+
  geom_bar(stat = "identity")+
  geom_errorbar(aes(ymin=(m.rmse)-(sd.rmse), 
                    ymax=(m.rmse)+(sd.rmse))) +
  plot.list +
  labs(y='averaged RMSE') ->p.rmse
p.rmse

# Combine R2 and RMSE plots in one figure 
p.r2 + p.rmse+ 
  plot_annotation(title="Goodness-of-fit measures", tag_levels=c("A"))

# save combined figure
ggsave("results/Figure8.Averaged_goodness.of.fit.measures.5cv.tiff", device="tiff", dpi = 300, height = 5, width = 7, limitsize = T )
```
