---
title: "Distributions of variables in Theoretical Models"
author: "Milou Sep"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    toc: yes
    number_sections: yes
    df_print: kable
    fig_width: 10
---

# Set-up: load data
```{r setup, results='hide', message=FALSE}
rm(list=ls())
```

```{r data, results='hide', message=FALSE}
source("R/SAMexplore_prepare.for.analyses.r")
```


# Untransformed variables (i.e. before z-scores)

## Covariates 

### Age
```{r untransformed age, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist2[[x]]$age, main = paste0('imputation ',x))}
```

### BMI 
```{r untransformed bmi, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist2[[x]]$BMI, main = paste0('imputation ',x))}
```

## Independent variables

### Childhood trauma
```{r untransformed CTQ, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist2[[x]]$JTV_total, main = paste0('imputation ',x))}
```

### Life-time stress
```{r untransformed LSCR, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist2[[x]]$LSC_event_total, main = paste0('imputation ',x))}
```

### Trait Anxiety
```{r untransformed TA, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist2[[x]]$STAI_T_total, main = paste0('imputation ',x))}
```

### AUCi sAA
```{r untransformed saa, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist2[[x]]$AUCi_sAA_1, main = paste0('imputation ',x))}
```

### AUCi Cortisol
```{r untransformed cort, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist2[[x]]$AUCi_CORT_1, main = paste0('imputation ',x))}
```

## Dependent variables

### Emotional memory
```{r untransformed emotional memory, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist2[[x]]$emotional.memory, main = paste0('imputation ',x))}
```

### Neutral memory
```{r untransformed neutral memory, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist2[[x]]$neutral.memory, main = paste0('imputation ',x))}
```

### Fear memory
```{r untransformed fear memory, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist2[[x]]$fear.memory, main = paste0('imputation ',x))}
```


# Standardized variables (i.e. after z-score transformations)

## Covariates 

### Age
```{r standardized age, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist3[[x]]$age, main = paste0('imputation ',x))}
```

### BMI 
```{r standardized bmi, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist3[[x]]$BMI, main = paste0('imputation ',x))}
```

## Independent variables

### Childhood trauma
```{r standardized CTQ, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist3[[x]]$JTV_total, main = paste0('imputation ',x))}
```

### Life-time stress
```{r standardized LSCR, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist3[[x]]$LSC_event_total, main = paste0('imputation ',x))}
```

### Cumulative childhood + life-time stress
```{r CTQ + LSCR, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist3[[x]]$stress_exposure, main = paste0('imputation ',x))}
```

### Trait Anxiety
```{r standardized TA, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist3[[x]]$STAI_T_total, main = paste0('imputation ',x))}
```

### AUCi sAA
```{r standardized saa, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist3[[x]]$AUCi_sAA_1, main = paste0('imputation ',x))}
```

### AUCi Cortisol
```{r standardized cort, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist3[[x]]$AUCi_CORT_1, main = paste0('imputation ',x))}
```

## Dependent variables

### Emotional memory
```{r standardized emotional memory, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist3[[x]]$emotional.memory, main = paste0('imputation ',x))}
```

### Neutral memory
```{r standardized neutral memory, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist3[[x]]$neutral.memory, main = paste0('imputation ',x))}
```

### Fear memory
```{r standardized fear memory, echo=FALSE}
par(mfrow = c(2, 5))
for(x in 1:M){hist(implist3[[x]]$fear.memory, main = paste0('imputation ',x))}
```


# Considerations for Box-Cox transformation

After inspection of the plots, the possibilities for a box-cox transformation were explored. Decided to only perform a z-transformation for all variables in TM (and RF), and not use box-cox or other transformations for TM, for the following reasons:

1. Residuals in TM are normal with z-transformations
2. Only age and childhood trauma have skewed distributions, other variables have normal distributions.
  + Childhood trauma is not in TM (only the cumulative score of childhood trauma and life-time stress), and age is a covariate.
  + transforming (only some) variables will only complicate interpretation
3. [It is no problem for linear regression if independent parameters are skewed](https://www.researchgate.net/post/Skewed_data_for_regression_analysis)
4. RF/Boruta is used next to the TM's to deal with non-linearity
5. Regular BC transformation is not possible for variables with negative values (like the relative memory outcomes in our data), a more complicated 2 parameters BC transformation would be needed..

```{r box-cox, eval=FALSE, include=FALSE}
abs(implist2[[1]]$neutral.memory)
 lm(neutral.memory~1+age+BMI+JTV_total*STAI_T_total*AUCi_sAA_1*AUCi_CORT_1, data=implist2[[1]])->model1
 lm(JTV_total~1+age+BMI, data=implist2[[1]])->model1

 plot(model1)

library(MASS)
bc=boxcox(model1, lambda = seq(-3,3))  # error door negative values
bc$x

library(geoR)
# 2 parameter box-cox? (for neg values?) https://www.researchgate.net/post/How_can_I_transform_a_data_series_with_negative_zero_and_positive_data_points
boxcoxfit(implist2[[1]]$JTV_total)->bc2
bc2
plot(bc2)

boxcoxfit(implist2[[1]]$STAI_T_total,lambda2 = T) ->bc3
bc3
bc3%>% plot

boxcoxfit(implist3[[1]]$stress_exposure, lambda = seq(-3,3), lambda2 = T , hist=T) ->bc4
plot(bc4)

 library(car)
 powerTransform(model1, family = 'bcnPower')
 powerTransform(model1, family = 'bcnPower', gamma= implist2[[1]]$subject)
```
