---
title: "Visualize TM results"
author: "Milou Sep"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    toc: yes
    number_sections: yes
    df_print: kable
editor_options: 
  chunk_output_type: console
---

# Preparations

The effects are visualized using the [estimated marginal effects](https://strengejacke.github.io/ggeffects/).

- Marginal effects measure the change in the conditional mean of outcome ð‘¦ when regressors change by one unit."https://stats.stackexchange.com/tags/marginal-effect/info"
- It is customary to present such marginal effects
+ evaluated at own values of ð‘¥ and ð‘§ and averaged for all individuals in the sample (average marginal effect or AME)
+ evaluated at the mean/median/modal values of ð‘¥ and ð‘§ (marginal effect at representative values or MER)
+ estimated at specific values that are interesting to the analyst

additional information on ggeffects

- https://rstudio-pubs-static.s3.amazonaws.com/490532_5b04406f4d154f2d9b5a99ff488d644a.html
- https://cran.r-project.org/web/packages/ggeffects/vignettes/ggeffects.html


First the required packages are added.
```{r setup, include=FALSE}
rm(list=ls())
# for imputed data
library(mice)
library(mitools)
#visualization & export
library(ggplot2)
library(ggpubr)
library(ggeffects) # for plotting interaction effects with EME
library(RColorBrewer)
# To make publication ready tables for word | info: https://davidgohel.github.io/flextable/articles/overview.html
library(flextable) 
library(officer)
```

```{r data, include=FALSE}
source("R/SAMexplore_prepare.for.analyses.r")
imputationList(imp.nos)->imp.nos2
imputationList(imp.ims)->imp.ims2
imputationList(imp.des)->imp.des2
```

general layout for plots
```{r}
my.plot.layout<-list( 
  theme_classic() ,
  theme(legend.position = "top"),
  scale_color_brewer(palette="Dark2"),
  theme(text = element_text(size=9)))
```


# Characteristics that influence memory context-dependency before acute stress

## Context-dependency of emotional information
There was a 3-way TA * cort * sAA interaction effect on context-dependency of emotional information before acute stress.

```{r EME TA * cort * sAA on emotional memory without acute stress}
fit.NS.EM.3way<-lm(emotional.memory~1+
                     age+BMI+stress_exposure+STAI_T_total+AUCi_sAA_1+AUCi_CORT_1+ #main effects
                     stress_exposure:STAI_T_total+ # 2 way
                     stress_exposure:AUCi_sAA_1+
                     stress_exposure:AUCi_CORT_1+
                     STAI_T_total:AUCi_sAA_1+
                     STAI_T_total:AUCi_CORT_1 +
                     STAI_T_total:AUCi_sAA_1:AUCi_CORT_1+ 
                     stress_exposure:AUCi_sAA_1:AUCi_CORT_1,
                   data = imp.nos[[10]])

# plots to explore the 3-way interaction
emm.NS.EM.3way <- ggpredict(fit.NS.EM.3way, c("AUCi_sAA_1", "AUCi_CORT_1",  "STAI_T_total"))

plot(emm.NS.EM.3way)
# adjust labels
levels(emm.NS.EM.3way$facet)<-c("STAI-T: mean-1SD", "STAI-T: mean","STAI-T: mean+1SD")
levels(emm.NS.EM.3way$group)<-c("mean-1SD", "mean","mean+1SD")

plot.NS.EM.TA.sAA.CO<-ggplot(emm.NS.EM.3way, aes(x, predicted, colour=group)) +
  geom_line() +
  facet_wrap(~facet) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .05, linetype='dotted',size=.3) + # remove lines with 'linetype=0'
  labs( title=parse(text = paste0('"Emotional MC ~ trait anxiety"','%*%', '"cortisol"', '%*%', '"\u03b1-amylase"')),
        subtitle="no stress (basal)",
        colour = bquote("Z "["cortisol (AUCi)"]),
        caption= "dPrime",
        x=bquote("Z "["\u03b1-amylase (AUCi)"]),
        y='predicted MC (EME)')+
  my.plot.layout
plot.NS.EM.TA.sAA.CO
```



# Characteristics that influence memory context-dependency 30min after acute stress
Visualizations of significant relations in immediate stress-group

## Context-dependency of emotional information
Adding cumulative stress exposure to the TM model without 2-way interactions significantly improved the model.

```{r EMM life-time stress on emotional memory 30min after acute stress}
fit.IS.EM.main<-lm(emotional.memory ~1+age+BMI+
                     stress_exposure+
                     STAI_T_total+
                     AUCi_sAA_1+
                     AUCi_CORT_1, 
                   data = imp.ims[[10]])

emm.IS.EM.CS <- ggpredict(fit.IS.EM.main, c( "stress_exposure"))

plot.IS.EM.CS<-ggplot(emm.IS.EM.CS, aes(x, predicted)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .1) +
  labs( title="Emotional MC ~ life adversity",
        subtitle="acute phase",
        caption="dPrime",
        x=bquote("Z "["CTQ"]* " + " * "Z " ["LSC-R"]),
        y='predicted MC (EME)')+
  my.plot.layout
plot.IS.EM.CS
```


## Context-dependency of neutral information
Adding AUCi of cortisol during learning to the TM model without 2-way interactions significantly improved the model.
```{r EMM cortisol on neutral memory 30min after acute stress}
fit.IS.NM.main<-lm(neutral.memory ~1+age+BMI+
                     stress_exposure+
                     STAI_T_total+
                     AUCi_sAA_1+
                     AUCi_CORT_1, 
                   data = imp.ims[[10]])

emm.IS.NM.CO <- ggpredict(fit.IS.NM.main, c( "AUCi_CORT_1"))

plot.IS.NM.CO<-ggplot(emm.IS.NM.CO, aes(x, predicted)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .1) +
  labs( title="Neutral MC ~ cortisol ",
        subtitle="acute phase",
        caption="dPrime",
        x=bquote("Z "["cortisol (AUCi)"]),
        y='predicted MC (EME)')+
  my.plot.layout
plot.IS.NM.CO
```

## Context-dependency of fear information
Two interactions improved the TM without 3-way interactions. [- no effects on fear learning after stress]

- stress-exposure x sAA (learning)
- trait anxiety x cortisol (learning)

### Interaction I: life adversity x AUC sAA

Note The interaction is visualized by calculated EMMs for mean(stress-exposure)-1sd(stress-exposure), mean(stress-exposure) and  mean(stress-exposure)+1sd(stress-exposure)
```{r}
imp.ims[[10]]$stress_exposure %>% summary()
imp.ims[[10]]$stress_exposure %>% mean()->mean
imp.ims[[10]]$stress_exposure %>% mean_sd()
imp.ims[[10]]$stress_exposure %>% sd()->sd
mean+sd
mean-sd
```


```{r EMM life-time stress * sAA on fear memory 30min after acute stress}
fit.IS.FM.2way<-lm(fear.memory~1+ age+BMI+stress_exposure+STAI_T_total+AUCi_sAA_1+AUCi_CORT_1+ #main effects
                     stress_exposure:STAI_T_total+ # 2 way
                     stress_exposure:AUCi_sAA_1+
                     stress_exposure:AUCi_CORT_1+
                     STAI_T_total:AUCi_sAA_1 +
                     STAI_T_total:AUCi_CORT_1,
                   data = imp.ims[[10]])

emm.IS.FM.CS.sAA <- ggpredict(fit.IS.FM.2way, terms=c( "AUCi_sAA_1", "stress_exposure"))

# adjust labels
levels(emm.IS.FM.CS.sAA$group)<-c("mean -1SD", "mean","mean +1SD")

plot.IS.FM.CS.sAA<-ggplot(emm.IS.FM.CS.sAA, aes(x, predicted, colour=group)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .05, linetype='dotted',size=.3) + # remove lines with 'linetype=0'
  labs( title=parse(text = paste0('"Fearful MC ~ life adversity"', '%*%', '"\u03b1-amylase"')),
        subtitle="acute phase",
        colour = bquote("Z "["CTQ"]* " + " * "Z " ["LSC-R"]),
        caption = "FPS",
        x=bquote("Z "["\u03b1-amylase (AUCi)"]),# unicode alpha code: https://stackoverflow.com/questions/5293715/how-to-use-greek-symbols-in-ggplot2
        y='predicted MC (EME)') +
  my.plot.layout
plot.IS.FM.CS.sAA
```


### Interaction II: Trait Anxiety x AUC cortisol (immediate stress-group)

```{r EMM TA * cortisol on fear memory 30min after acute stress}
emm.IS.FM.TA.CO <- ggpredict(fit.IS.FM.2way, terms=c( "AUCi_CORT_1", "STAI_T_total"))

# adjust labels
levels(emm.IS.FM.TA.CO$group)<-c("mean-1SD", "mean","mean+1SD")

plot.IS.FM.TA.CO<-ggplot(emm.IS.FM.TA.CO, aes(x, predicted, colour=group)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .05, linetype='dotted',size=.3) +
  labs( title=parse(text = paste0('"Fearful MC ~ trait anxiety"', '%*%', '"cortisol"')),
        subtitle="acute phase",
        colour = bquote("Z "["STAI-T"]),
        caption = "FPS",
        x=bquote("Z "["cortisol (AUCi)"]),
        y='predicted MC (EME)')+
  my.plot.layout
plot.IS.FM.TA.CO
```


# Characteristics that influence memory context-dependency 2h after acute stress

## Context-dependency of neutral information

Two interactions improved the TM without 3-way interactions.

- stress-exposure x sAA (learning)
- stress-exposure x cortisol (learning)

### Interaction I: life adversity x AUC sAA
more stress exposure, less negative effect of sAA on neutral memory contextualization?

```{r EMMs life-time stress * sAA on neutral memory 2h after acute stress}
fit.DS.NM.2way<-lm(neutral.memory~1+ age+BMI+stress_exposure+STAI_T_total+AUCi_sAA_1+AUCi_CORT_1+ #main effects
                     stress_exposure:STAI_T_total+ # 2 way
                     stress_exposure:AUCi_sAA_1+
                     stress_exposure:AUCi_CORT_1+
                     STAI_T_total:AUCi_sAA_1 +
                     STAI_T_total:AUCi_CORT_1,
                   data = imp.des[[10]])

emm.DS.NM.CS.sAA <- ggpredict(fit.DS.NM.2way, terms=c( "AUCi_sAA_1", "stress_exposure"))

# adjust labels
levels(emm.DS.NM.CS.sAA$group)<-c("mean-1SD", "mean","mean+1SD")

plot.DS.NM.CS.sAA<-ggplot(emm.DS.NM.CS.sAA, aes(x, predicted, colour=group)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .05, linetype='dotted',size=.3) +
  labs( title=parse(text = paste0('"Neutral MC ~ life adversity"', '%*%', '"\u03b1-amylase"')),
        subtitle="recovery phase",
        colour = bquote("Z "["CTQ"]* " + " * "Z " ["LSC-R"]),
        caption="dPrime",
        x=bquote("Z "["\u03b1-amylase (AUCi)"]),
        y="predicted MC (EME)")+
  my.plot.layout
plot.DS.NM.CS.sAA
```

### Interaction II: Cumulative stress exposure and Cortisol
```{r EMMs life-time stress * cortisol on neutral memory 2h after acute stress}
emm.DS.NM.CS.CO <- ggpredict(fit.DS.NM.2way, terms=c( "AUCi_CORT_1", "stress_exposure"))

# adjust labels
levels(emm.DS.NM.CS.CO$group)<-c("mean-1SD", "mean","mean+1SD")

plot.DS.NM.CS.CO<-ggplot(emm.DS.NM.CS.CO, aes(x, predicted, colour=group)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .05, linetype='dotted',size=.3) +
  labs( title=parse(text = paste0('"Neutral MC ~ life adversity"', '%*%', '"cortisol"')),
        subtitle="recovery phase",
        colour = bquote("Z "["CTQ"]* " + " * "Z " ["LSC-R"]),
        caption="dPrime",
        x=bquote("Z "["cortisol (AUCi)"]),
        y="predicted MC (EME)")+
  my.plot.layout
plot.DS.NM.CS.CO
```


# Summary & save

All effects (in manuscript)
```{r}
ggarrange(
  # immediate
  ggarrange(
    plot.IS.EM.CS, 
    plot.IS.NM.CO,
    plot.IS.FM.CS.sAA,
    plot.IS.FM.TA.CO,
    ncol = 2, nrow = 2,
    heights=c(1,1.5),
    labels=c("A", "B", "C", "D"),  common.legend = F),
  
  # delayed
  ggarrange( 
    plot.DS.NM.CS.sAA,
    plot.DS.NM.CS.CO,
    ncol = 2, nrow = 1,
    labels = c("E","F"),
    common.legend = F),
  
  # no stress
  plot.NS.EM.TA.sAA.CO,
  ncol = 1, nrow = 3, 
  heights=c(1.5,1,1),
  labels=c("","","G") 
)

ggsave("results/Figure.4.TM.tiff", device="tiff", dpi = 300, height = 12, width = 8, limitsize = T )
```
