---
title: "Descriptive Tabels SAMexplore"
author: "Milou Sep"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    toc: yes
    number_sections: yes
    df_print: kable
    fig_width: 10
---

# Set-up: load data
```{r setup}
rm(list=ls())
library(arsenal) # https://cran.r-project.org/web/packages/arsenal/vignettes/tableby.html
library(dplyr)
```


this data was created by reprocessing raw data (with prep code in imputation script), this preprocessed data was then saved to the implist en auc were calculated in AUCcalc script. THEN use part of'prepare for analysis code to select only vars included in data
```{r load data}
readRDS("processed_data/implist.withRAW11.rds")->implist
M=11
implist[[11]] ->raw_df
implist[[1]] ->imp_df
rbind(raw_df, imp_df)->merged_df
```

# Descriptives missing values
```{r}
# before imputation
# Number of missing values per composite variable (before multiple imputation). 
# https://cran.r-project.org/web/packages/arsenal/vignettes/tableby.html#modify-how-missing-values-are-displayed
apply(is.na(raw_df),2,sum) %>% data.frame() ->missings # 2 indicates columns
which(rownames(missings) == "fear.learning")
summary(missings[-which(rownames(missings) == "fear.learning"),]) #NB fear.at learning not included in further analysis.
missings[-which(rownames(missings) == "fear.learning"),]->miss.raw
min(miss.raw)/nrow(raw_df) *100
max(miss.raw)/nrow(raw_df) *100
mean(miss.raw)/nrow(raw_df) *100

# The number of participants with at least one missing value (before multiple imputation) 
raw_df %>% select(!fear.learning) -> dat1
apply(is.na(dat1),1,sum)  ->missings.per.participant # 1 indicates rows.
sum(missings.per.participant > 1)
(sum(missings.per.participant > 1) / length(missings.per.participant)) *100

# after imputation
# Number of missing values per composite variable (after multiple imputation)
apply(is.na(imp_df),2,sum) %>% data.frame() ->missings.imp
missings.imp[-which(rownames(missings.imp) == "fear.learning"),]->miss.imp
summary(missings.imp[-which(rownames(missings.imp) == "fear.learning"),])
min(miss.imp)/nrow(imp_df) *100
max(miss.imp)/nrow(imp_df) *100
(mean(miss.imp)/nrow(imp_df)) *100

# The number of participants with at least one missing value (after multiple imputation) 
imp_df %>% select(!fear.learning) -> dat2
apply(is.na(dat2),1,sum)  ->missings.per.participant.after.imputation # 1 indicates rows.
sum(missings.per.participant.after.imputation > 1) 
(sum(missings.per.participant.after.imputation > 1) / length(missings.per.participant.after.imputation)) *100
```

# Collect variabels that were included in boruta
```{r}
# calculate z-scores [Note, same formula as in 'prepare for analysis' script]
raw_df %>% 
  # https://www.r-bloggers.com/r-tutorial-series-centering-variables-and-generating-z-scores-with-the-scale-function/
  mutate_if(is.numeric, ~as.vector(scale(.,center = TRUE, scale = TRUE))) ->x.new2 # solution to drop attributes (from scale)

# create cumulative stress measure by sum of centered z-scores:
raw_df$stress_exposure<- x.new2$JTV_total + x.new2$LSC_event_total

plyr::revalue(raw_df$Condition, c("No-Stress"="No stress group", "Immediate-Stress"="Acute phase group", "Delayed-Stress" = "Recovery phase group"))->raw_df$Condition

# Change labels to English
mutate_at(raw_df,c("roken", "alcohol", "drugs", "nachtritme_verst"), ~recode(.,"[0] Nee"="No", "[1] Ja"="Yes"))->raw_df
mutate_at(raw_df,c("psychol", "lich_ziekte"), ~recode(.,"0"="No", "1"="Yes"))->raw_df
```

# Relabel variables
```{r}
table.labels<-c(subject="subject ID",
                BMI	= "BMI",
                age	= "age",
                Condition =	"Experimental Group",
                roken	= "smoking",
                alcohol	= "alcohol",
                drugs	= "recreational drugs",
                psychol	= "history of mental illnesses",
                lich_ziekte	= "history of physical illnesses",
                nachtritme_verst =	"disrupted day/night rhythm",
                SCL_total	= "SCL (total)",
                JTV_total	= "CTQ (total)",
                LSC_event_total =	"LSC-R (total)",
                stress_exposure	= "Life Adversity (z(CTQ )+ z(LSCR))",
                
                STAI_T_total =	"STAI-T (total)",
                HEXACO_hh.mean = "Honesty-Humility (HEXACO subscale)",         
                HEXACO_emo.mean = "Emotionality (HEXACO subscale)",
                HEXACO_extr.mean = "Extraversion (HEXACO subscale)",       
                HEXACO_ag.mean = "Agreeableness (HEXACO subscale)",
                HEXACO_co.mean = "Conscientiousness (HEXACO subscale)",
                HEXACO_op.mean = "Openness to Experience (HEXACO subscale)",
                
                VTCI_NS_total = 'Novelty Seeking (TCI-SF subscale)', 
                VTCI_HA_total= 'Harm Avoidance (TCI-SF subscale)', 
                VTCI_RD_total= 'Reward Dependence (TCI-SF subscale)', 
                VTCI_PS_total= 'Persistence (TCI-SF subscale)', 
                VTCI_SD_total= 'Self-Directedness (TCI-SF subscale)', 
                VTCI_CO_total= 'Cooperativeness (TCI-SF subscale)', 
                VTCI_ST_total= 'Self-Transcendence (TCI-SF subscale)',
                
                AUCg_CORT_1 =	"Cortisol (AUCg) at learning",
                AUCg_sAA_1	= "sAA (AUCg) at learning",
                AUCg_CORT_2	= "Cortisol (AUCg) at retention",
                AUCg_sAA_2	= "sAA (AUCg) at retention",
                AUCi_CORT_1	= "Cortisol (AUCi) at learning",
                AUCi_sAA_1	= "sAA (AUCi) at learning",
                AUCi_CORT_2	= "Cortisol (AUCi) at retention",
                AUCi_sAA_2	= "sAA (AUCi) at retention",

                AUCg_VAS_anxiety_1	= "VAS arousal (AUCg) at learning",
                AUCg_VAS_mood_1	= "VAs mood (AUCg) at learning",
                AUCg_VAS_anxiety_2	= "VAS arousal (AUCg) at retention",
                AUCg_VAS_mood_2	= "VAS mood (AUCg) at retention",
                AUCi_VAS_anxiety_1 =	"VAS arousal (AUCi) at learning",
                AUCi_VAS_mood_1	= "VAS mood (AUCi) at learning",
                AUCi_VAS_anxiety_2 =	"VAS arousal (AUCi) at retention",
                AUCi_VAS_mood_2 =	"VAS mood (AUCi) at retention",

                AUCg_stais_1 =	'STAI-S (AUCg) at learning',
                AUCg_stais_2 =	"STAI-S (AUCg) at retention",
                AUCi_stais_1 =	"STAI-S (AUCi) at learning",
                AUCi_stais_2 =	"STAI-S (AUCi) at retention",
                AUCg_panas_NegA_1	= "PANAS Negative Affect (AUCg) at learning",
                AUCg_panas_PosA_1	= "PANAS Positive Affect (AUCg) at learning",
                AUCg_panas_NegA_2	= "PANAS Negative Affect (AUCg) at retention",
                AUCg_panas_PosA_2	= "PANAS Positive Affect (AUCg) at retention",
                AUCi_panas_NegA_1	= "PANAS Negative Affect (AUCi) at learning",
                AUCi_panas_PosA_1	= "PANAS Positive Affect (AUCi) at learning",
                AUCi_panas_NegA_2	= "PANAS Negative Affect (AUCi) at retention",
                AUCi_panas_PosA_2	= "PANAS Positive Affect (AUCi) at retention",
                
                fear.memory = "Fearful MC",
                fear.learning = "Only for fear MC: fear acquisition",
                neutral.memory = "Neutral MC",
                emotional.memory = "Emotional MC"
)

!(data.frame(table.labels) %>% row.names() %in% 
    c("fear.memory", #"fear.at learning", 
      "neutral.memory", "emotional.memory", "Condition")) ->boruta.variabels.index

# export labeled variables
data.frame(table.labels)[boruta.variabels.index,] %>% write.csv2(.,"results/boruta.vars.labeled.csv")
```

# set "layout" controls tableby 
```{r}
# https://rdrr.io/cran/arsenal/man/tableby.control.html
mycontrols  <- tableby.control( test=T, total=T,
                                na.rm=F,
                                numeric.stats=c( "meansd","range","Nmiss"),
                                cat.stats = c("countpct", "Nmiss"),
                                ord.stats= c("countpct","Nmiss"),
                                stats.labels=list(Nmiss='Missing', Nmiss2='Missing', countpct="test",
                                                  meansd="Mean (SD)", range="Range"),
                                pfootnote=TRUE,
                                cat.simplify = T, numeric.simplify = T, ordered.simplify = T,
                                digits=2,
                                digits.p=2,
                                digits.count=0,
                                digits.pct=2 )
```

# Create table
```{r}
raw_df %>% select(-c(fear.learning, subject)) ->raw_df

# create table (of raw data)
tableby(Condition ~., data=raw_df, control = mycontrols)->raw.table

summary(raw.table,
        title = "Raw scores SAM study",
        labelTranslations = table.labels)
# tests(raw.table)
```

# Export to word file
```{r}
# save tables to word
write2word(raw.table, "results/descriptive table (raw).doc", title="Raw scores SAM study",
           labelTranslations = table.labels)
# write2word(merged.table, "results/descriptive table (raw and imp).doc", title="Raw and imputed scores SAM study")
```

# Exprot to CSV file
```{r}
# write.csv2(as.data.frame(summary(raw.table)),"results/descriptive table (raw).csv")
```
