---
title: "Variable selction based on variable importance in random forests"
author: "Milou Sep"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    toc: yes
    number_sections: yes
    df_print: kable
    fig_width: 10
---

# Background

*"Boruta is an all relevant feature selection wrapper algorithm, capable of working with any classification method that output variable importance measure (VIM); by default, Boruta uses Random Forest. The method performs a top-down search for relevant features by comparing original attributes' importance with importance achievable at random, estimated using their permuted copies, and progressively eliminating irrelevant features to stabilise that test."*[boruta r-docu](https://www.rdocumentation.org/packages/Boruta/versions/6.0.0/topics/Boruta)

further reading on boruta: https://www.analyticsvidhya.com/blog/2016/03/select-important-variables-boruta-package/
[boruta package](https://www.datacamp.com/community/tutorials/feature-selection-R-boruta)
[useful boruta review](https://watermark.silverchair.com/bbx124.pdf?token=AQECAHi208BE49Ooan9kkhW_Ercy7Dm3ZL_9Cf3qfKAc485ysgAAAm0wggJpBgkqhkiG9w0BBwagggJaMIICVgIBADCCAk8GCSqGSIb3DQEHATAeBglghkgBZQMEAS4wEQQMMzKPDZoDXuNoe6pWAgEQgIICIPG8O1WRq6zMO-AN-xPBUoWaDnTclVVVefwA6rPH_wGCAKBL4BUCehQeIsyBDAACFohULG6GWk5CHhF2DzVKv0nl51_cSzpZHSgNGvw8XN93cxFKc6UddXCFhdvanqcZjDK06MRwEBbQ0qUHr_afGuV5MlkJbEUIlvYfk8JzYCvucNT0W3HPtfPRlPjA8hCHo-GoKx9IGi1pArc5HHNNA-zP5OZQaHZnI8QZLwbpwj6StmgoCSSy5Yla1yforU3HBFymhTOasqF_CUaf3_xEKn-iWY9hy_TUlrfT6i14IaXlkRxUNJWIoaF3ms8nJPcM3dt3LSMRx69K-T8PPJ90dwff43iq-0FVAs4iy7i7eVWOJ_5OAavrzhFcWzFIFts1_g9wdt7QZgULU_NxRPIEj8SjREyEHujk3uGCMwrPt9yUKPQOHOsckuco69uW9yscJAMBTM1r1PCNmuGD5mkqg9lXOtEuaB5OSE2A91h5jyrO2BVFfso6G2_lheey0ncsrE3zXle9DFd55FAwqeg24skDXDp5fHCsKKzrVk3N_-IpBPDaVeESmrim2z_exp6UqFgjQ4vAb79CKk4oCdft8bscPwIccUsPkk_AIQBlx2Y88ph3WyYrJJD0x7Y6X_gJDnydA1mAy_ls-FSEyQTxT2WO34kIXCrT1vz85Su4knSZlhQFLrqOWwijiuhKnbUx4V9Eug8n2NbegidCV_hpf1Q])

# Set-up
required packages and (preprocessed and imputed) data
```{r setup, results='hide', message=FALSE}
rm(list=ls())
# load packages
library(Boruta)
# for export tables
library(flextable)
library(officer)

library(knitr)
library(kableExtra)
```

load imputed data, prepared for SAM_explore analysis RFs / Boruta / lm's
```{r data, results='hide', message=FALSE}
source("R/SAMexplore_prepare.for.analyses.r")
seed=29180
```


# Tuning parameters Boruta

## RF tuning
To get an idea of the optimal tuning parameters for randomForest, the RMSE (with 5-fold cross-validation) of the models with parameters: 1) .mtry=c(2:56), and 2) .ntree=c(500, 1000, 1500, 2000, 2500) were checked on the 10th imputed data set of each experimental group (no-stress, delayed-stress, immediate-stress), for all memory types (fearful, neutral, emotional information). The Rmarkdown file "Tune_Boruta.Rmd" contains the code for these checks.

In all experimental groups, for all memory types mtry=2 resulted in the lowest RMSE. For ntree optimal numbers were between 500-1500. Since an increase number of tree's will only improve performance 1500 was selected as minimum

## adapting RF settings for Boruta
Based on the tuning results we started with mtry=2 and ntree=1500. Boruta feature selection with these parameters, resulted in some tentative results (neither confirmed, nor rejected). We tested the effect of an increased ntree from 1500 to 2000 and maxrun from 1000 to 1500, however this did not affect the number of tentative variables. The number of tentative variables was in general very low and not present in all imputed datasets of a certain memory type or experimental group. 

In this document, the tentative values are displayed in the variable selection plots from in each imputed dataset. However, we decided to use the TentativeRoughFix() function to make a decision on the tentative variables (features with a median importance higher then the maximal one of the the shadow attributes were confirmed others rejected) for majority voting.

```{r boruta parameters}
m.try=2
n.tree=1500
max.r=1000
```


# Step 1: Boruta Feature selection
The Boruta feature selection is applied to each imputed dataset of an experimental group and Boruta objects are plotted via function below.
```{r plot boruta}
plot.boruta<-function(boruta_object){
  plot(boruta_object, xlab = "", xaxt = "n")
  lz<-lapply(1:ncol(boruta_object$ImpHistory),function(i)
    boruta_object$ImpHistory[is.finite(boruta_object$ImpHistory[,i]),i])
  names(lz) <- colnames(boruta_object$ImpHistory)
  Labels <- sort(sapply(lz,median))
  axis(side = 1,las=2,labels = names(Labels),
       at = 1:ncol(boruta_object$ImpHistory), cex.axis = 0.7)
}
```

## Boruta fear memory
```{r boruta function fear memory}
boruta.fear<-function(imp_list){
  boruta.F<-list()
  for(i in 1:M){        
    prep.fear.m(imp_list[[i]])->RFfear
    set.seed(seed)
    boruta.fear <- Boruta(fear.memory~., data = RFfear, doTrace = 0, maxRun=max.r, mtry=m.try, ntree=n.tree)
    # print results with tentatives
    print(boruta.fear) #if there are tentative vars increase maxRuns (default = 100)
    # lm(getConfirmedFormula(boruta.fear_train),RFfear)->fit1
    # anova(fit1)
    # boruta.fear_train$finalDecision
    plot.boruta(boruta.fear)->plot
    print(plot)
    # fix tentative
    TentativeRoughFix(boruta.fear)->boruta.fear2
    #extract attributes (=results)
    attStats(boruta.fear2)->results
    results$varname<-rownames(results)
    #bind results to list
    results->boruta.F[[i]]
  }
  return(boruta.F)
}
```

### No-stress group: fear memory
```{r boruta fear memory no-stress}
boruta.fear(imp.nos)->n_f
saveRDS(n_f,"processed_data/n_f.RDS")
# readRDS("processed_data/n_f.RDS")->n_f
```

### Immediate stress group: fear memory
```{r boruta fear memory immediate stress}
boruta.fear(imp.ims)->i_f
saveRDS(i_f,"processed_data/i_f.RDS")
# readRDS('processed_data/i_f.RDS')->i_f
```

### Delayed stress group: fear memory
```{r boruta fear memory delayed stress}
boruta.fear(imp.des)->d_f
saveRDS(d_f,"processed_data/d_f.RDS")
# readRDS('processed_data/d_f.RDS')->d_f
```

## Boruta emotional memory
```{r boruta function emotional memory}
boruta.emo<-function(imp_list){
  boruta.E<-list()
  for(i in 1:M){        
    prep.emo(imp_list[[i]])->RFemo
    set.seed(seed)
    boruta.emo <- Boruta(emotional.memory~., data =RFemo , doTrace = 0, maxRun=max.r, mtry=m.try, ntree=n.tree)
    # print results with tentatives
    print(boruta.emo)
    plot.boruta(boruta.emo)->plot
    print(plot)
    # fix tentative
    TentativeRoughFix(boruta.emo)->boruta.emo2
    #extract attributes (=results)
    attStats(boruta.emo2)->results
    #add rownames as variable
    results$varname<-rownames(results)
    #bind results to list
    results->boruta.E[[i]] 
  }
  return(boruta.E)
}
```

### No-stress group: emotional memory
```{r boruta emotional memory no-stress}
boruta.emo(imp.nos)->n_e
saveRDS(n_e,'processed_data/n_e.RDS')
# readRDS('processed_data/n_e.RDS')->n_e
```

### Immediate stress group: emotional memory
```{r boruta emotional memory immediate stress}
boruta.emo(imp.ims)->i_e
saveRDS(i_e,'processed_data/i_e.RDS')
# readRDS('processed_data/i_e.RDS')->i_e
```

### Delayed stress group: emotional memory
```{r boruta emotional memory delayed stress}
boruta.emo(imp.des)->d_e
saveRDS(d_e,'processed_data/d_e.RDS')
# readRDS('processed_data/d_e.RDS')->d_e
```

## Boruta neutral memory
```{r boruta function neutral memory}
boruta.neu<-function(imp_list){
  boruta.N<-list()
  for(i in 1:M){        
    prep.neu(imp_list[[i]])->RFneu
    set.seed(seed)
    boruta.neu <- Boruta(neutral.memory~., data =RFneu , doTrace = 0, maxRun=max.r, mtry=m.try, ntree=n.tree)
    # print results with tentatives
    print(boruta.neu)
    plot.boruta(boruta.neu)->plot
    print(plot)
    # fix tentative
    TentativeRoughFix(boruta.neu)->boruta.neu
    #extract attributes (=results)
    attStats(boruta.neu)->results
    #add rownames as variable
    results$varname<-rownames(results)
    #bind results to list
    results->boruta.N[[i]]
  }
  return(boruta.N)
}
```

### No-stress group: neutral memory
```{r boruta neutral memory no-stress}
boruta.neu(imp.nos)->n_n
saveRDS(n_n,'processed_data/n_n.RDS')
# readRDS('processed_data/n_n.RDS')->n_n
```

### Immediate stress group: neutral memory
```{r boruta neutral memory immediate stress}
boruta.neu(imp.ims)->i_n
saveRDS(i_n,'processed_data/i_n.RDS')
# readRDS('processed_data/i_n.RDS')->i_n
```

### Delayed stress group: neutral memory
```{r boruta neutral memory delayed stress}
boruta.neu(imp.des)->d_n
saveRDS(d_n,'processed_data/d_n.RDS')
# readRDS('processed_data/d_n.RDS')->d_n
```


# Step 2: Majority Votes (>=6)
For each memory type are for each experimental group 10 matrices with boruta results. Of each matrix the variable names with their final decisions are extracted.
```{r}
calc_majority_vote<-function(results_list){
  #create empty list
  selected_list<-list()
  #select varnames and decision from each imputed datasets
  for(i in 1:M){
    results_list[[i]] %>% select(varname, decision) ->selected_list[[i]]
  }
  #merge selected data from imputations in one dataframe by variable names
  plyr::join_all(selected_list, by='varname')->df_majority
  # add unique variable names
  colnames(df_majority)<-c("varname", paste0("imp.",rep(1:10)))
  # add column that indicate if a variable is more than 5 times confirmed
  df_majority$majority_vote<- (rowSums(df_majority[,-1] == 'Confirmed')>5)
  #return dataframe
  return(df_majority)
}
```

## No-stress group
### Fear memory
```{r}
calc_majority_vote(n_f)->n_f_majority
n_f_majority %>% filter(majority_vote == TRUE)
n_f_majority %>% filter(majority_vote == TRUE) %>% select(varname)->no.stres.fear
```

### Emotional memory
```{r}
calc_majority_vote(n_e)->n_e_majority
n_e_majority %>% filter(majority_vote == TRUE)
n_e_majority %>% filter(majority_vote == TRUE) %>% select(varname)->no.stres.emo
```

### Neutral memory
```{r}
calc_majority_vote(n_n)->n_n_majority
n_n_majority %>% filter(majority_vote == TRUE) 
n_n_majority %>% filter(majority_vote == TRUE) %>% select(varname)->no.stres.neu
```


## Immediate stress group
### Fear memory
```{r}
calc_majority_vote(i_f)->i_f_majority
i_f_majority %>% filter(majority_vote == TRUE)
i_f_majority %>% filter(majority_vote == TRUE) %>% select(varname)->im.stres.fear
```

### Emotional memory
```{r}
calc_majority_vote(i_e)->i_e_majority
i_e_majority %>% filter(majority_vote == TRUE)
i_e_majority %>% filter(majority_vote == TRUE) %>% select(varname)->im.stres.emo
```

### Neutral memory
```{r}
calc_majority_vote(i_n)->i_n_majority
i_n_majority %>% filter(majority_vote == TRUE) 
i_n_majority %>% filter(majority_vote == TRUE) %>% select(varname)->im.stres.neu
```


## Delayed stress group
### Fear memory
```{r}
calc_majority_vote(d_f)->d_f_majority
d_f_majority %>% filter(majority_vote == TRUE)
d_f_majority %>% filter(majority_vote == TRUE) %>% select(varname)->de.stres.fear
```

### Emotional memory
```{r}
calc_majority_vote(d_e)->d_e_majority
d_e_majority %>% filter(majority_vote == TRUE)
d_e_majority %>% filter(majority_vote == TRUE) %>% select(varname)->de.stres.emo
```

### Neutral memory
```{r}
calc_majority_vote(d_n)->d_n_majority
d_n_majority %>% filter(majority_vote == TRUE) 
d_n_majority %>% filter(majority_vote == TRUE) %>% select(varname)->de.stres.neu
```


# Step 3: Summary tables

```{r echo=FALSE}
results.no.s <- qpcR:::cbind.na(no.stres.fear, no.stres.emo, no.stres.neu)
colnames(results.no.s) <- c( "Fear",  "Emotional", "Neutral")
results.im.s <- qpcR:::cbind.na(im.stres.fear, im.stres.emo, im.stres.neu)
colnames(results.im.s) <- c( "Fear",  "Emotional", "Neutral")
results.de.s <- qpcR:::cbind.na(de.stres.fear, de.stres.emo, de.stres.neu)
colnames(results.de.s) <- c( "Fear",  "Emotional", "Neutral")
```

Kable tables
```{r, echo=FALSE, message=FALSE}
results.no.s %>%
  kable(caption="No stress group") %>%
  kable_styling()

results.im.s %>%
  kable(caption="Immedidate stress group") %>%
  kable_styling()

results.de.s %>%
  kable(caption="Delayed stress group") %>%
  kable_styling()
```

Export to Microsoft Word
```{r echo=TRUE, results='hide'}
# Use flextable layout to prepare table (for word)
flextable(results.no.s) %>% #set_caption(., "No stress group", html_escape = T) %>% 
  add_header_lines(., values = c("No stress group") ) %>%
  colformat_char(.,na_str='')%>%
  theme_vanilla(.)%>% autofit(.)->results.ns

flextable(results.im.s) %>% #set_caption(., "Immediate stress group", html_escape = T) %>% 
  add_header_lines(., values = c("Immedidate stress group") ) %>%
  colformat_char(.,na_str='')%>%
  theme_vanilla(.)%>% autofit(.)->results.is  

flextable(results.de.s) %>% #set_caption(., "Delayed stress group", html_escape = T) %>% 
  add_header_lines(., values = c("Delayed stress group") ) %>%
  colformat_char(.,na_str='')%>%
  theme_vanilla(.)%>% autofit(.)->results.ds  

# export data to word   
word_export <- read_docx()
# body_add_par(word_export, value = "No stress group")
body_add_flextable(word_export,  value = results.ns, align="left")  # https://stackoverflow.com/questions/49150272/exporting-several-tables-to-word-documents-using-r-package-flextable
body_add_par(word_export, value = "")
# body_add_par(word_export, value = "Immediate stress group")
body_add_flextable(word_export, results.is, align="left")
body_add_par(word_export, value = "")
# body_add_par(word_export, value = "Delayed stress group")
body_add_flextable(word_export, results.ds, align="left")
print(word_export, 'results/Boruta_results.docx')
```
